--- 
title: "Analyzing Shifts"
author: "Sam Gasell"
date: "1/30/2021"
output: html_document
---

Read in data generated from Situational Shifts.Rmd

```{r}
library(tidyverse)

#Replace with own file path
totalTeamStats1920 <- read_csv("/Users/samgasell/Documents/Hockey/Shift Stats by Team 19-20.csv")

#Combine 6on5 and 5on6 data with 5on4 and 4on5, respectively. This is because many teams do not have enough shifts in 4on5 or 5on4 situations, so combines to categorize when teams have a 1 skater advantage or deficit.
totalTeamStats1920 <- totalTeamStats1920 %>% mutate(totalTime1Up = totalTime6o5 + totalTime5o4) %>%     mutate(totalTime1Down = totalTime5o6 + totalTime4o5) %>% mutate(t1Up = t6o5 + t5o4) %>% mutate(t1Down = t5o6 + t4o5) %>% mutate(s1Up = s6o5 + s5o4) %>% mutate(s1Down = s5o6 + s4o5) %>% mutate(GA1Up = GA6o5 + GA5o4) %>% mutate(GA1Down = GA5o6 + GA4o5) %>% mutate(GF1Up = GF6o5 + GF5o4) %>% mutate(GF1Down = GF5o6 + GF4o5) %>% mutate(a1Up = t1Up/s1Up) %>% mutate(a1Down = t1Down/s1Down)
```

Regression and plots representing the Poisson regression between average shift time and offensive scoring rate in each manpower situation.

6-on-6 -- Significant
```{r}
#Poisson regression
poiFit6o6GF <- glm(GF6o6 ~ offset(log(totalTime6o6/(60*60))) + a6o6, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a6o6, y = log(GF6o6/(totalTime6o6/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Scored per 60 Minutes by Team", subtitle = "6-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a6o6, y = GF6o6/(totalTime6o6/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit6o6GF$coefficients[1] + poiFit6o6GF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "6-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit6o6GF)
```
5-on-5 
```{r}
#Poisson regression
poiFit5o5GF <- glm(GF5o5 ~ offset(log(totalTime5o5/(60*60))) + a5o5, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a5o5, y = log(GF5o5/(totalTime5o5/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Scored per 60 Minutes by Team", subtitle = "5-on-5 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a5o5, y = GF5o5/(totalTime5o5/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit5o5GF$coefficients[1] + poiFit5o5GF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "5-on-5 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit5o5GF)
```
4-on-4
```{r}
#Poisson regression
poiFit4o4GF <- glm(GF4o4 ~ offset(log(totalTime4o4/(60*60))) + a4o4, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a4o4, y = log(GF4o4/(totalTime4o4/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Scored per 60 Minutes by Team", subtitle = "4-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a4o4, y = GF4o4/(totalTime4o4/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit4o4GF$coefficients[1] + poiFit4o4GF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "4-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit4o4GF)
```

1 Player Advantage
```{r}
#Poisson regression
poiFit1UpGF <- glm(GF1Up ~ offset(log(totalTime1Up/(60*60))) + a1Up, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a1Up, y = log(GF1Up/(totalTime1Up/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Scored per 60 Minutes by Team", subtitle = "1 Player Advantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a1Up, y = GF1Up/(totalTime1Up/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit1UpGF$coefficients[1] + poiFit1UpGF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "1 Player Advantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit1UpGF)
```

6-on-4
```{r}
#Poisson regression
poiFit6o4GF <- glm(GF6o4 ~ offset(log(totalTime6o4/(60*60))) + a6o4, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a6o4, y = log(GF6o4/(totalTime6o4/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Scored per 60 Minutes by Team", subtitle = "6-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a6o4, y = GF6o4/(totalTime6o4/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit6o4GF$coefficients[1] + poiFit6o4GF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "6-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit6o4GF)
```

1 Player Disadvantage -- Significant
```{r}
#Poisson regression
poiFit1DownGF <- glm(GF1Down ~ offset(log(totalTime1Down/(60*60))) + a1Down, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a1Down, y = log(GF1Down/(totalTime1Down/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "1 Player Disadvantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a1Down, y = GF1Down/(totalTime1Down/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit1DownGF$coefficients[1] + poiFit1DownGF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "1 Player Disadvantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit1DownGF)
```


4-on-6
```{r}
#Poisson regression
poiFit4o6GF <- glm(GF4o6 ~ offset(log(totalTime4o6/(60*60))) + a4o6, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a4o6, y = log(GF4o6/(totalTime4o6/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Scored per 60 Minutes by Team", subtitle = "4-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Scored/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a4o6, y = GF4o6/(totalTime4o6/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit4o6GF$coefficients[1] + poiFit4o6GF$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Scored per 60 Minutes by Team", subtitle = "4-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Scored/60 Minutes")

#Summary of regression
summary(poiFit4o6GF)
```


Goals Allowed

6-on-6
```{r}
#Poisson regression
poiFit6o6GA <- glm(GA6o6 ~ offset(log(totalTime6o6/(60*60))) + a6o6, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a6o6, y = log(GA6o6/(totalTime6o6/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Allowed per 60 Minutes by Team", subtitle = "6-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a6o6, y = GA6o6/(totalTime6o6/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit6o6GA$coefficients[1] + poiFit6o6GA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "6-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit6o6GA)
```
5-on-5 
```{r}
#Poisson regression
poiFit5o5GA <- glm(GA5o5 ~ offset(log(totalTime5o5/(60*60))) + a5o5, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a5o5, y = log(GA5o5/(totalTime5o5/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Allowed per 60 Minutes by Team", subtitle = "5-on-5 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a5o5, y = GA5o5/(totalTime5o5/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit5o5GA$coefficients[1] + poiFit5o5GA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "5-on-5 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit5o5GA)
```
4-on-4
```{r}
#Poisson regression
poiFit4o4GA <- glm(GA4o4 ~ offset(log(totalTime4o4/(60*60))) + a4o4, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a4o4, y = log(GA4o4/(totalTime4o4/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Allowed per 60 Minutes by Team", subtitle = "4-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a4o4, y = GA4o4/(totalTime4o4/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit4o4GA$coefficients[1] + poiFit4o4GA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "4-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit4o4GA)
```

1 Player Advantage - Significant
```{r}
#Poisson regression
poiFit1UpGA <- glm(GA1Up ~ offset(log(totalTime1Up/(60*60))) + a1Up, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a1Up, y = log(GA1Up/(totalTime1Up/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Allowed per 60 Minutes by Team", subtitle = "1 Player Advantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a1Up, y = GA1Up/(totalTime1Up/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit1UpGA$coefficients[1] + poiFit1UpGA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "1 Player Advantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit1UpGA)
```

6-on-4
```{r}
#Poisson regression
poiFit6o4GA <- glm(GA6o4 ~ offset(log(totalTime6o4/(60*60))) + a6o4, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a6o4, y = log(GA6o4/(totalTime6o4/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Allowed per 60 Minutes by Team", subtitle = "6-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a6o4, y = GA6o4/(totalTime6o4/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit6o4GA$coefficients[1] + poiFit6o4GA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "6-on-4 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit6o4GA)
```

1 Player Disadvantage
```{r}
#Poisson regression
poiFit1DownGA <- glm(GA1Down ~ offset(log(totalTime1Down/(60*60))) + a1Down, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a1Down, y = log(GA1Down/(totalTime1Down/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "1 Player Disadvantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a1Down, y = GA1Down/(totalTime1Down/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit1DownGA$coefficients[1] + poiFit1DownGA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "1 Player Disadvantage, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit1DownGA)
```


4-on-6 - Significant
```{r}
#Poisson regression
poiFit4o6GA <- glm(GA4o6 ~ offset(log(totalTime4o6/(60*60))) + a4o6, data = totalTeamStats1920, family = "poisson")

#Plot of log of scoring rate versus average shift time with linear regression
ggplot(totalTeamStats1920, aes(x = a4o6, y = log(GA4o6/(totalTime4o6/(60*60))), label=teamAbb)) +  geom_label() + geom_smooth(method = 'lm', se = FALSE) + labs(title="Average Shift Time vs. Logarithm of Goals Allowed per 60 Minutes by Team", subtitle = "4-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "log(Goals Allowed/60 Minutes)")

#Plot of the scoring rate versus average shift time with exponential regression
ggplot(totalTeamStats1920, aes(x = a4o6, y = GA4o6/(totalTime4o6/(60*60)), label=teamAbb)) +  geom_label() + stat_function(fun = function(x) exp(poiFit4o6GA$coefficients[1] + poiFit4o6GA$coefficients[2]*x), color = 'blue', size = 1.3) + labs(title="Average Shift Time vs. Goals Allowed per 60 Minutes by Team", subtitle = "4-on-6 Manpower, 2019-2020 NHL Regular season", x = "Average Shift Time (Seconds)", y = "Goals Allowed/60 Minutes")

#Summary of regression
summary(poiFit4o6GA)
```
